\documentclass{article}

\usepackage{../../../../doc/latex/cactus}

\begin{document}

\title{The {\tt Llama} multipatch infrastructure}
\author{Original authors: Denis Pollney, Christian Reisswig, Erik Schnetter, Nils Dorband, Peter Diener}

\date{\today}

\maketitle

% START CACTUS THORNGUIDE

\begin{abstract}
  {\tt Llama} is a collection of thorns that defines a set of curvilinear multipatch grids in Cactus.
  The grids are allowed to overlap. Interpatch boundary data is exchanged via interpolation.
\end{abstract}

\section{Introduction}
\label{sec:intro}

The {\tt Llama}\footnote{The name Llama originated from our affinity to the animal (not really).
Others remember the computer game {\tt SimCity 2000}, where Llamas regularly appeared throughout the game.
Llamas are funny creatures.}
code was implemented to allow one to use multiple curvilinear grid patches within {\tt Cactus}.
The ability to have multiple grid ``maps'' is provided by {\tt Carpet}. 
Patch systems, and their coordinate systems and transformations are defined in {\tt Llama}.

The core of {\tt Llama} consists of the following core thorns:
\begin{itemize}
 \item {\tt Coordinates}: Provides patch systems, coordinate mappings, and transformations.
 \item {\tt Interpolate2}: Sets up inter-patch interpolation. The interpolation itself is carried out using {\tt CarpetInterp2}.
\end{itemize}

In addition, there are a number of optional thorns:
\begin{itemize}
 \item {\tt GlobalDerivatives}: Provides inlined pointwise finite difference operators in the global Cartesian tensor basis. 
       The differencing is carried out in the local basis. Jacobians are applied to transform to the global basis.
\end{itemize}




\section{Using Llama}
\label{sec:use}

What follows is a brief introduction to using {\tt Llama}. It assumes that
you are familiar with the two papers describing Llama \cite{Llama_Pollney:2009yz,Llama_Reisswig:2012nc}.

{\tt Llama: Coordinates} provides grid functions for Jacobians, inverse Jacobians, and the derivitives
of the Jacobian between the local coordinates of a given patch and the global Cartesian coordinates.
It also provides a volume form to compute volume integrals in the global frame, taking into account the non-trivial
overlap between patches.


\subsection{Obtaining This Thorn}

The public version of {\tt Llama} can be found on the
website {\tt http://www.llama-code.org}. 

\subsection{Basic Usage}

% Describe basic parameters: overlap, interpolation order/type, ghostsize width, 
% storage for Jacobians / volume form.

Here, we describe the basic parameter settings for {\tt Coordinates} and {\tt Interpolate2}.
Parameters that are specific to a given patch system are described in the sections below.

The following parameters must be set for multipatch evolution:
\begin{itemize}
 \item {\verb|Carpet::domain_from_multipatch|} must be set to {\tt yes} for multipatch.
  \item {\verb|CartGrid3D::type|} must be set to {\tt multipatch} for multipatch.
  \item {\verb|CartGrid3D::set_coordinate_ranges_on|} must be set to {\tt "all maps"} for multipatch.
\end{itemize}

The following basic parameters need to be set 
 \begin{itemize}
 \item {\verb|Coordinates::coordinate_system|}. This selects a particular patch system defined in Llama.
 \item {\verb|Coordinates::patch_boundary_size|}. Sets the ghost size of the interpatch boundary. This depends, e.g., on the finite difference stencil that is used.
       This parameter should be equal to {\verb|Driver::ghost_size|}.
 \item {\verb|Coordinates::additional_overlap_size|}. Sets the number of additional overlap points. The size of this parameter depends on the interpolation order and type that is used,
       and on the patch system. One wants this parameters to be as small as possible. {\tt Llama} will complain if the overlap size is not sufficient for a given patch system
       and interpolation setup.
 \item {\verb|Coordinates::outer_boundary_size|}. Sets the number of ghost zones at \textit{outer} boundaries (i.e., ghost data is not obtained from inter-patch interpolation)
       where evolution specific outer boundary conditions must be applied.
 \item {\verb|Coordinates::store_jacobian|}. Activate/deactivate storage for Jacobian grid function. Whether storage is required depends on other thorns that may use this function.
 \item {\verb|Coordinates::store_inverse_jacobian|}. Activate/deactive storage for inverse Jacobian grid function. Whether storage is required depends on other thorns that may use this function.
 \item {\verb|Coordinates::store_volume_form|}. Activate/deactive storage for volume form grid function. Whether storage is required depends on other thorns that may use this function.
 \item {\verb|Interpolate::interpolator_order|}. Selects the order of Lagrange inter-patch interpolation.
 \item {\verb|Interpolate::interpolator_order_matter|}. Selects the order of ENO interpolation for those variables that contain the ``matter'' tag.
       If the default value {\tt -1} is used, this tag will be ignored and Lagrange interpolation is used also for ``matter'' variables.
\end{itemize}
  
With cell-centered AMR, the following parameters must be set:
\begin{itemize}
 \item {\verb|Coordinates::stagger_outer_boundaries = yes|}.
 \item {\verb|Coordinates::stagger_patch_boundaries = yes|}.
\end{itemize}


\subsection{Patch systems}

Currently, we support the following patch systems:
\begin{itemize}
 \item {\tt Cartesian}: This is standard Cartesian coordinates.
 \item {\tt TwoPatchCartesian}: This is two Cartesian patches, with one common face.
 \item {\tt TwoPatchDistorted}: This is one Cartesian patch and a second patch with distortion. It is useful for testing code.
 \item {\tt Thornburg04}: This is a 7-patch system consisting of 1 central Cartesian cube capable of AMR which is surrounded by 6 ``inflated-cube'' spherical grids. 
 \item {\tt Thornburg04nc}: This is a 6-patch system suitable for excision. It consists of 6 ``inflated-cube'' spherical grids which together cover a spherical shell.
 \item {\tt Thornburg13}: This is similar to the 7 patch system {\tt Thornburg04}, but with 13 patches.
 \item {\tt CylinderInBox}: This is a hollow (spherical) cylinder in a (Cartesian) box.
 \item {\tt Sphere+Column}: This is another system suitable for excision. It has an excision-type overlapping sphere and column grid.
 \item {\tt Cylinder+Column}: This is similar to the {\tt Sphere+Column} system, but with a central cylindrical grid instead of a spherical one.
\end{itemize}

\subsubsection{Thornburg04: 7-patch system}

The following basic parameters need to be set for this patch system: 
 \begin{itemize}
 \item {\verb|Coordinates::sphere_outer_radius|}. The physical radius of the outer boundary defined by the ``inflated-cube'' spherical grids.
 \item {\verb|Coordinates::sphere_inner_radius|}. The radius of the inner boundary of the ``inflated-cube'' spherical grids to the central cartesian cube.
       The size of the central Cartesian cube is controlled by this parameter.
 \item {\verb|Coordinates::h_cartesian|}. The coarse grid resolution $\Delta x$ of the central Cartesian patch.
 \item {\verb|Coordinates::h_radial|}. The radial resolution $\Delta r$ of the ``inflated-cube'' spherical grids.
 \item {\verb|Coordinates::n_angular|}. The number of grid cells per angular direction per patch of the ``inflated-cube'' spherical grids.
 \item {\verb|Coordinates::radial_stretch|}. Use radial stretching on the spherical inflated-cube grids {\tt yes/no}.
 \item {\verb|Coordinates::stretch_rmin_1|}. Radius from which we start to decrease radial resolution.
 \item {\verb|Coordinates::stretch_rmax_1|}. Radius at which radial resolution becomes {\verb|Coordinates::h_radial_1|} and remains constant again
 \item {\verb|Coordinates::h_radial_1|}. Radial stretched target resolution.
\end{itemize}

\subsubsection{Thornburg04nc: 6-patch system}

This has the same parameters as the Thornburg04 patch system, with the
exception of {\verb|Coordinates::h_cartesian|}, which has no meaning.

%\subsubsection{13-patch system}


\subsection{Visualizing multi-patch data}
The \texttt{CarpetHDF5} plugin for \texttt{VisIt}~\cite{Llama_VisIt:web}
supports reading multi-patch
HDF5 files out of the box. It does however require that coordinates were
output along with the actual data files. Assuming that your current options
for HDF5 output look like this
\begin{verbatim}
IOHDF5::out_vars = "HydroBase::rho"
\end{verbatim}
it is sufficient to change them to
\begin{verbatim}
CarpetIOHDF5::one_file_per_group = "no" # this is required by multipatch
IOHDF5::out_vars = "HydroBase::rho grid::coordinates"
\end{verbatim}
\texttt{VisIt} will present the Cartesian and curvilinear parts of the grid as
two different meshes and two different variables.


% \subsection{Support and Feedback}



\section{History}

The approximate time line is something like this:
\begin{itemize}
\item ~2009: First version of Llama: Ability to simulate vacuum binary black hole mergers.
\item ~2012: Second version of Llama: Ability to simulate general-relativistic hydrodynamics.
\end{itemize}


% \subsection{Thorn Source Code}


% \subsection{Thorn Documentation}


% \subsection{Acknowledgements}

\begin{thebibliography}{20}
%\cite{Pollney:2009yz}
\bibitem{Llama_Pollney:2009yz}
  D.~Pollney, C.~Reisswig, E.~Schnetter, N.~Dorband and P.~Diener,
  %``High accuracy binary black hole simulations with an extended wave zone,''
  Phys.\ Rev.\ D {\bf 83}, 044045 (2011), arXiv:0910.3803.
  %%CITATION = doi:10.1103/PhysRevD.83.044045;%%
  %82 citations counted in INSPIRE as of 22 Nov 2016

%\cite{Reisswig:2012nc}
\bibitem{Llama_Reisswig:2012nc}
  C.~Reisswig, R.~Haas, C.~D.~Ott, E.~Abdikamalov, P.~Mösta, D.~Pollney and E.~Schnetter,
  %``Three-Dimensional General-Relativistic Hydrodynamic Simulations of Binary Neutron Star Coalescence and Stellar Collapse with Multipatch Grids,''
  Phys.\ Rev.\ D {\bf 87}, 064023 (2013), arXiv:1212.1191.
  %%CITATION = doi:10.1103/PhysRevD.87.064023;%%
  %20 citations counted in INSPIRE as of 22 Nov 2016

\bibitem{Llama_VisIt:web}
{VisIt} visualization tool \url{https://wci.llnl.gov/codes/visit/}

\end{thebibliography}

% END CACTUS THORNGUIDE

\end{document}
