\documentclass{article}

% Use the Cactus ThornGuide style file
% (Automatically used from Cactus distribution, if you have a
%  thorn without the Cactus Flesh download this from the Cactus
%  homepage at www.cactuscode.org)
\usepackage{../../../../../doc/latex/cactus}

\newlength{\tableWidth} \newlength{\maxVarWidth} \newlength{\paraWidth} \newlength{\descWidth} \begin{document}

\title{PUGHInterp}
\author{Paul Walker, Thomas Radke, Erik Schnetter}
\date{$ $Date$ $}

\maketitle

% Do not delete next line
% START CACTUS THORNGUIDE

\newcommand{\InterpGridArrays}{{\tt CCTK\_\-Interp\-Grid\-Arrays()}}
\newcommand{\PUGHInterp}{{\it PUGH\-Interp}}

\begin{abstract}
Thorn \PUGHInterp\ implements the Cactus interpolation API \InterpGridArrays\
for the interpolation of CCTK grid arrays at arbitrary points.
\end{abstract}

\section{Introduction}
Thorn \PUGHInterp\ provides an implementation of the Cactus interpolation API
specification for the interpolation of CCTK grid arrays at arbitrary points,
\InterpGridArrays.

This function interpolates a list of CCTK grid arrays (in a multiprocessor run
these are generally distributed over processors) on a list of interpolation
points. The grid topology and coordinates are implicitly specified via a Cactus
coordinate system.
The interpolation points may be anywhere in the global Cactus grid.
In a multiprocessor run they may vary from processor to processor;
each processor will get whatever interpolated data it asks for.

The routine \InterpGridArrays\ does not do the actual interpolation
itself but rather takes care of whatever interprocessor communication may be
necessary, and -- for each processor's local patch of the domain-decomposed grid
arrays -- calls {\tt CCTK\_InterpLocalUniform()} to invoke an external
local interpolation operator (as identified by an interpolation handle).
It is advantageous to interpolate a list of grid arrays at once (for the same
list of interpolation points) rather than calling \InterpGridArrays\ several
times with a single grid array. This way note only can \PUGHInterp's
implementation of \InterpGridArrays\ aggregate communications for multiple grid
arrays into one (resulting in less communications overhead) but also
{\tt CCTK\_InterpLocalUniform()} may compute interpolation coefficients once
and reuse them for all grid arrays.

Please refer to the {\it Cactus UsersGuide} for a complete function description
of \InterpGridArrays\ and {\tt CCTK\_InterpLocalUniform()}.\\


\section{\PUGHInterp's Implementation of \InterpGridArrays}

If thorn \PUGHInterp\ was activated in the {\tt ActiveThorns} list of a
parameter file for a Cactus run, it will overload at startup the flesh-provided
dummy function for \InterpGridArrays\ with its own routine. This routine will
then be invoked in subsequent calls to \InterpGridArrays.

\PUGHInterp's routine for the interpolation of grid arrays provides exactly the
same semantics as \InterpGridArrays which is thoroughly described in the
{\it Function Reference} chapter of the {\it Cactus UsersGuide}.
In the following, only user-relevant details about its implementation, such as
specific error codes and the evaluation of parameter options table entries, are
explained.

\subsection{Implementation Notes}

At first, \InterpGridArrays\ checks its function arguments for invalid values
passed by the caller. In case of an error, the routine will issue an error
message and return with an error code of either {\tt UTIL\_\-ERROR\_\-BAD\_\-HANDLE} for
an invalid coordinate system and/or parameter options table, or
{\tt UTIL\_ERROR\_BAD\_INPUT} otherwise.
Currently there is the restriction that only {\tt CCTK\_VARIABLE\_REAL} is
accepted as the CCTK data type for the interpolation points coordinates.

Then the parameter options table is parsed and evaluated for additional
information about the interpolation call (see section \ref{PUGHInterp_PTable}
for details).

In the single-processor case, \InterpGridArrays\ would now invoke the local
interpolation operator (as specified by its handle) by a call to
{\tt CCTK\_InterpLocalUniform()} to perform the actual interpolation. The return
code from this call is then also passed back to the user.

For the multi-processor case, \PUGHInterp\ does a query call to the local
interpolator first to find out whether it can deal with the number of
interprocessor ghostzones available. For that purpose it sets up an array of
two interpolation points which denote the extremes of the physical coordinates
on a processor: the lower-left and upper-right point of the processor-local
grid's bounding box\footnote{
Note that because the query is done with extreme interpolation points
coordinates, the interpolation call may fail even if all the user-supplied
interpolation points are well within each processor's local patch.
The reason for this implementation behaviour is that we safely want to catch
all errors caused by a too small ghostzone size.}.
The query gets passed the same user-supplied function arguments as for the
real interpolation call, apart from the interpolation points coordinates (which
now describe a processor's physical bounding box coordinates) and the output
array pointers (which are all set to NULL in order to indicate that this is a
query call only). A return code of \verb|CCTK_ERROR_INTERP_POINT_OUTSIDE|
from \verb|CCTK_InterpLocalUniform()| for this query call (meaning the local interpolator potentially requires values
from grid points which are outside of the available processor-local patch of
the global grid) causes \InterpGridArrays\ to return immediately with
a \verb|CCTK_ERROR_INTERP_GHOST_SIZE_TOO_SMALL| error code on all processors.

Otherwise the \InterpGridArrays\ routine will continue and map the user-supplied
interpolation points onto the processors which own these points. In a subsequent
global communication all processors receive "their" interpolation points
coordinates and call {\tt CCTK\_InterpLocalUniform()} with those. The
interpolation results are then sent back to the processors which originally
requested the interpolation points.

Like the {\it PUGH} driver thorn, \PUGHInterp\ uses MPI for the necessary
interprocessor communication. Note that the {\tt MPI\_Alltoall()/MPI\_Alltoallv()} calls for the
distribution of interpolation points coordinates to their owning processors and
the back transfer of the interpolation results to the requesting processors
are collective communication operations. So in the multi-processor case you
{\em must\/} call \InterpGridArrays\ in parallel on {\em each\/} processor
(even if a processor doesn't request any points to interpolate at), otherwise
the program will run into a deadlock.


\subsection{Passing Additional Information via the Parameter Table}
\label{PUGHInterp_PTable}

One of the function arguments to \InterpGridArrays\ is an integer handle
which refers to a key/value options table. Such a table can be used to pass
additional information (such as the interpolation order) to the interpolation
routines (i.e.\  to both \InterpGridArrays\ and the local interpolator as invoked
via {\tt CCTK\_InterpLocalUniform()}). The table may also be modified by these
routines, eg. to exchange internal information between the local and global
interpolator, and/or to pass back arbitrary information to the user.

The only table option currently evaluated by \PUGHInterp's implementation
of \InterpGridArrays\ is:
\begin{verbatim}
  CCTK_INT input_array_time_levels[N_input_arrays];
\end{verbatim}
which lets you choose the timelevels for the individual grid arrays to
interpolate (in the range $[0, N\_time\_levels\-\_of\_var\_i - 1]$).
If no such table option is given, then the current timelevel (0)
will be taken as the default.

The following table options are meant for the user to specify how the
local interpolator should deal with interpolation points near grid boundaries:
\begin{verbatim}
  CCTK_INT  N_boundary_points_to_omit[2 * N_dims];
  CCTK_REAL boundary_off_centering_tolerance[2 * N_dims];
  CCTK_REAL boundary_extrapolation_tolerance[2 * N_dims];
\end{verbatim}
In the multi-processor case, \InterpGridArrays\ will modify these arrays in
a user-supplied options table in order to specify the handling of interpolation
points near interprocessor boundaries (ghostzones) for the local interpolator;
corresponding elements in the options arrays are set to zero for all ghostzone
faces, i.e.\  no points should be omitted, and no off-centering and extrapolation
is allowed at those boundaries. Array elements for physical grid boundaries are
left unchanged by \InterpGridArrays.

If any of the above three boundary handling table options is missing in the
user-supplied table, \InterpGridArrays\ will create and add it to the table
with appropriate defaults. For the default values, as well as a comprehensive
discussion of grid boundary handling options, please refer to documentation
of the thorn(s) providing local interpolator(s) (eg. thorn {\it LocalInterp} in
the {\it Cactus ThornGuide}).

At present, the table option \verb|boundary_extrapolation_tolerance| is not
implemented. Instead, if any point cannot be mapped onto a processor
(i.e.\  the point is outside the global grid), a level-1 warning is printed to
stdout by default, and the error code \verb|CCTK_ERROR_INTERP_POINT_OUTSIDE| is
returned. The warning will not be printed if the parameter table contains an
entry (of any type) with the key \verb|"suppress_warnings"|.

The local interpolation status will be stored in the user-supplied parameter
table (if given) as an integer scalar value with the option key
\verb|"local_interpolator_status"| (see section \ref{PUGHInterp_return_codes} for details).

The table options
\begin{verbatim}
  CCTK_POINTER per_point_status;
  CCTK_INT     error_point_status;
\end{verbatim}
are used internally by \InterpGridArrays\ to pass information
about per-point status codes between the global and the local interpolator
(again see section \ref{PUGHInterp_return_codes} for details).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{\InterpGridArrays\ Return Codes}
\label{PUGHInterp_return_codes}

The return code from \InterpGridArrays%%%
\footnote{%%%
	 In C the return code is the \InterpGridArrays\
	 function result; in Fortran it's returned through
	 the first ({\tt status}) argument.%%%
	 }%%%
is determined as follows:
\begin{itemize}
\item	If any of the arguments are invalid (e.g. $\verb|N_dims| < 0$),
	the return code is \verb|UTIL_ERROR_BAD_INPUT|.
\item	If any errors are encountered when processing the
	parameter table, the return code is the
	appropriate \verb|UTIL_ERROR_TABLE_*| error code.
\item	If the query call determines that
	the number of ghost zones in the grid is too small
	for the local interpolator, the return code is
	\verb|CCTK_ERROR_INTERP_POINT_OUTSIDE|.
\item	Otherwise, the return code from \InterpGridArrays\ is
	the minimum over all processors of the return code
	from the local interpolation on that processor.
\end{itemize}

If the local interpolator supports per-point status returns
and the user supplies an interpolator parameter table,
then in addition to this global interpolation return code,
\InterpGridArrays\ also returns a ``local'' status code which
describes the outcome of the local interpolation for all the
interpolation points which originated on {\em this\/} processor:
\begin{verbatim}
  CCTK_INT local_interpolator_status;
\end{verbatim}
This gives the minimum over all the interpolation points originating
on {\em this\/} processor, of the \verb|CCTK_InterpLocalUniform()|
return codes for those points.  (It doesn't matter on which processor(s)
the points were actually interpolated -- \InterpGridArrays\ takes care
of gathering all the status information back to the originating processors.)

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Comments}
For more information on how to invoke interpolation operators please refer
to the flesh documentation.

% Do not delete next line
% END CACTUS THORNGUIDE



\section{Parameters} 


\parskip = 0pt
\parskip = 10pt 

\section{Interfaces} 


\parskip = 0pt

\vspace{3mm} \subsection*{General}

\noindent {\bf Implements}: 

interp
\vspace{2mm}

\vspace{5mm}

\noindent {\bf Provides}: 



DriverInterpolate to 
\vspace{2mm}\parskip = 10pt 

\section{Schedule} 


\parskip = 0pt


\noindent This section lists all the variables which are assigned storage by thorn CactusPUGH/PUGHInterp.  Storage can either last for the duration of the run ({\bf Always} means that if this thorn is activated storage will be assigned, {\bf Conditional} means that if this thorn is activated storage will be assigned for the duration of the run if some condition is met), or can be turned on for the duration of a schedule function.


\subsection*{Storage}NONE
\subsection*{Scheduled Functions}
\vspace{5mm}

\noindent {\bf CCTK\_STARTUP} 

\hspace{5mm} pughinterp\_startup 

\hspace{5mm}{\it pughinterp startup routine } 


\hspace{5mm}

 \begin{tabular*}{160mm}{cll} 
~ & After:  & driver\_startup \\ 
~ & Language:  & c \\ 
~ & Type:  & function \\ 
\end{tabular*} 



\vspace{5mm}\parskip = 10pt 
\end{document}
